\section{Background}
\label{sec:background}

Three emerging non-volatile memory technologies promise to provide fast,
 non-volatile, byte-addressable memories: spin-torque transfer RAM (STT-RAM),
phase change memory (PCM), and memristor-based memories. In the following
sections we briefly introduce each of them.

\textbf{Spin-Torque Transfer RAM} STT-RAM uses magnetic properties of 
certain materials and electric charges to enable different persistent states 
in memory cells. STT-RAM promises potential of unlimited write cycles, and
current prototype using tunneling magnetorresistence showed 32ns for
read latency and 40ns for write latency, with 0.3mA power consumption
per cell [??]. The disadvantage of STT-RAM is shrinking cell size is
extremely difficult.

\textbf{Memristor memory} is a resistive RAM device. Applying electric field
changes the resistence of the memristor device, and different resistence level
can be read as 0 or 1. Recent research on memristor indicates that its
performance could be comparable to DRAM, with write endurance up to $10^{10}$
cycles. Currently memristor is still under development in the laboratory.

\textbf{Phase Change Memory} is a promising non-volatile byte-addressable
memory technology. PCM is built with Chalcogenide-based materials. By
injecting current to heat the memory cell to a high temperature for a period
of time, the material crystallize and the resistence reduce. To reset the cell,
apply a large enough current for a short period to melt down the
material and it transition into amorphous state. These two states can be read as
"1" and "0". 

PCM has fast reads comparable to DRAM, while writes are slower than DRAM.
As heating causes contraction of the cell, PCM has limited write endurance
up to $10^8$ cycles.
 
\subsection{NVRAM on Background}
\label{sec:background}

